\vspace{1.0cm} \Large \bf Integrating influenza antigenic dynamics with molecular evolution

### 2013-07-24 response.tex

The incidence data are poorly explained. In the first paragraph of the Epidemiological consequences of antigenic drift section, the authors use relative incidence. In the next paragraph, are they switching to absolute incidence? This is not clear. If they are switching to absolute incidence, I don't think they ever explain in the Methods where that data comes from. In either case, what are the justifications for using absolute versus relative incidence? Also, why do they not analyze incidence prior to 1998/1999? Finally, because each influenza season spans two calendar years, why do they not break up the data by season rather than year?

An interesting result is that estimating the serum effects improves the performance a great deal (more than including the phylogenetic data, in fact!). It suggests that there may be systematic differences between the sera. Can this be attributed to the serum itself (such as whether it is goat versus ferret), or to the particular lab or study? Could the authors correlate the estimated serum effects back to the originating source for that serum to check?

\textbf{We would suggest that a serum with a larger effect contains a more concentrated and active set of antibodies than a serum with a smaller effect. The increased number of antibodies could be due to experimental variation in serum extraction and processing or due to variation in immune response and timing between ferrets, i.e. some ferrets may mount an immediate and strong immune response, while others may mount a weaker response.  This causes variation in overall strength of the serum.  We try to control for the overall strength when looking at patterns of cross-reactivity through the use of `serum effects'.  Some differences in serum effects could be tracked down to particular studies, but this work is beyond the scope of the current study.  Here, we've essentially treated serum effect as a nuisance parameter.  We've revised the text to discuss this as well as describe why estimating serum effects improves model performance over fixing serum effects.}

### 2013-07-24 flux.tex
Fix #3 - serum effects.

We further extend the model by estimating the strength of overall reactivity of each serum rather than fixing this at its maximum titer, and additionally, by estimating the strength of reactivity of each virus isolate.
We refer to these estimates as serum effects and virus effects, respectively.
These effects determine the baseline expectation of titer when virus and serum have identical antigenic locations.
We find that including these effects decreases test error further, resulting in comparable improvements from estimating virus and serum effects individually and giving an improvement of between 0.03 to 0.13 log$_2$ HI titers when estimating both virus and serum effects together (models 8--10; Table \ref{errortable}).

>>>>

We further extend the model by estimating `serum effects' to quantify the overall strength reactivity of each serum, rather than fixing serum effects at maximum titers.
Serum effects likely derive from experimental variation in serum extraction and processing or from variation in immune responses between ferrets.
Serum effects determine the baseline expectation of titer when virus and serum have identical antigenic locations.
However, if serum effect is fixed to the serum's maximum titer, this will often not be the case, as the virus giving the maximum titer is likely to be antigenically distinct from the serum.
Thus, fixing serum effects will tend to under-estimate effect size; we observe a mean effect of 10.42 log$_2$ HI titers for A/H3N2 when fixing serum effects and a mean of 10.94 when estimating serum effects.
We find that estimating serum effects improves test error further (model 8, Table \ref{errortable}), with improvements of between 0.01 and 0.12 log$_2$ HI titers.

Additionally, we include and estimate `virus effects' in an analogous fashion, which are intended to represent differences in overall reactivity between viruses.
With virus effects included, baseline titer derives from both the virus and the serum used in the HI reaction.
We find that including virus effects further improves test error, either alone with fixed serum effects (model 9, Table \ref{errortable}) or combined with estimating serum effects (model 10, Table \ref{errortable}).
With fixed serum effects, the inclusion of virus effects results in improvements of between 0.02 and 0.09 log$_2$ HI titers and with estimated serum effects, the inclusion of virus effects results in improvements of between 0.01 and 0.05 log$_2$ HI titers.
